---
title: "Computing Macroecological Patterns"
author: "XXXXXXXXXXXXXXXXXX"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
---

### load libraries

```{r, results= "hide"}
# Load necessary software packages
library(rmarkdown)  # For creating dynamic documents

# Data wrangling and analysis
library(tidyverse)  # Collection of packages for data manipulation and visualization
library(ade4)       # Analysis of Ecological Data: functions for multivariate analysis
library(inlmisc)    # Miscellaneous functions for Bayesian models
library(parallel)   # For parallel processing

# Data download
library(rgbif)      # Interface to the Global Biodiversity Information Facility API
library(countrycode) # Convert country names to/from ISO country codes

# Spatial data handling
library(sf)         # Simple Features for R: handling spatial data
library(terra)      # For working with raster data
library(maptools)   # Tools for reading and handling spatial objects


# database harmonisation
library(taxize)            # For taxonomic data retrieval and manipulation
library(AmphiNom)          # AmphiNom: Cleaning amphibian occurrence data

# Spatial data cleaning
library(passport)          # Tools for cleaning geographical coordinates
library(CoordinateCleaner) # Cleaning geographical coordinates and taxonomic names

# Species distribution modelling packages
library(biomod2)   # Species distribution modeling
library(ENMeval)   # Evaluation of Ecological Niche Models


```


This script was used to quantify the macroecological patterns of Amphibians. But the script is generally transferable to other taxa.

Examples of the output files produced from this script can be downloaded in XXXXXXXXXXXXXXXXX.

# Preparing data

## Range map

### Step 1 : Upload the range maps. 

Download rangemaps and store this in "./data/rangeMaps" folder. The range maps should ideally be a an simple features class (package "sf"). Range maps will be used in computing all macroecological patterns.
```{r}
rangeMaps <- st_read(choose.files())
```

### Step 2 : Harmonise the taxonomy of the species

```{r}
# pull the scientific names column
rangeMaps_spp <- rangeMaps %>% 
  pull(sci_name) %>% 
  unique()

# update the taxonomies using amphinom::harmoniseSync. This uses the Amphibian Species of the Wolrd taxonomies (https://amphibiansoftheworld.amnh.org/)
rangeMaps_spp_resolve <- harmoniseSync(query = rangeMaps_spp) %>% 
  distinct(query, .keep_all = TRUE) %>% 
  mutate(match_count = str_count(harmonise_names, "\\w+")) %>% 
  mutate(scientificName_harmonise = if_else(match_count == 2, harmonise_names, query)) %>% 
  rename("binomial" = "query") %>% 
  dplyr::select(binomial, scientificName_harmonise)

# update the "rangeMaps" shapefile by merging the harmonised taxonomies.
rangeMaps <- rangeMaps %>% 
  left_join(rangeMaps_spp_resolve,
            by = c("sci_name" = "binomial"))%>%    
  relocate(scientificName_harmonise)

```


### Step 3 : computing geographic range from occurrence data

But some species do not have range maps. So you have to produce range maps from occurrence data. 




```{r}
# Download the occurrence data from GBIF. Upload the occurrence data. Put the occurrence data in the "./data/occurrenceData/ folder.

occs <- read.delim(file = file.choose(),
                            header = TRUE,
                            sep = "\t") %>% 
  rename("scientificName" = "species")

# Clean the occurrence data.

occs <- occs %>%
  dplyr::select(scientificName,
                decimalLongitude,
                decimalLatitude,
                countryCode,
                individualCount,
                gbifID,
                family, 
                taxonRank, 
                coordinateUncertaintyInMeters, 
                year,
                basisOfRecord, 
                institutionCode) %>% 
  rename("decimallongitude" = "decimalLongitude",
         "decimallatitude" = "decimalLatitude") %>%
  mutate_all(na_if, "") %>% 
  mutate(decimallongitude = as.numeric(decimallongitude),
         decimallatitude = as.numeric(decimallatitude)) %>% 
  filter(!is.na(scientificName_harmonise),
         !is.na(decimallongitude),
         !is.na(decimallatitude)) %>% 
  cc_dupl()  %>% 
  cc_zero() %>% 
  cc_equ() %>% 
  cc_val() %>% 
  cc_sea() %>% 
  cc_cap(geod = FALSE, buffer = 0.0416) %>% 
  cc_cen(geod = FALSE, buffer = 0.0416) %>% 
  cc_gbif(geod = FALSE, buffer = 0.0416) %>% 
  cc_inst(geod = FALSE, buffer = 0.0416) %>% 
  filter(coordinateUncertaintyInMeters  < 5000  | is.na(coordinateUncertaintyInMeters))

# harmonise occurrence data
occs_spp <- occs %>% 
  pull(scientificName) %>% 
  unique()


temp <- harmoniseSync(query = occs_spp)
occs_spp_resolve <- temp %>% 
  distinct(query, .keep_all = TRUE) %>% 
  mutate(match_count = str_count(harmonise_names, "\\w+")) %>% 
  mutate(scientificName_harmonise = if_else(match_count == 2, harmonise_names, query)) %>% 
  rename("binomial" = "query") %>% 
  dplyr::select(binomial ,scientificName_harmonise)


occs <- occs %>% 
  left_join(occs_spp_resolve,
            by = c("scientificName" = "binomial"))%>%    
  relocate(scientificName_harmonise)


# additional range maps


temp_occs_data <- occs %>% 
  filter(!(scientificName_harmonise %in% pull(st_drop_geometry(geographicRange), scientificName_harmonise))) 
temp_spp <- temp_occs_data %>% 
  distinct(scientificName_harmonise) %>% 
  pull(scientificName_harmonise)

# upload world map

data(wrld_si)
wm <- ne_download(scale = "large",
                  type = "land",
                  category = "physical",
                  returnclass = "sf") %>% st_transform(crs = "+proj=longlat +datum=WGS84")

# now make additional range maps

i <- temp_spp[1]
temp_occs <- temp_occs_data %>% 
      filter(scientificName_harmonise == i) %>% 
      st_as_sf(coords = c("decimallongitude", "decimallatitude"),
                          crs = "+proj=longlat +datum=WGS84")
temp_range <- st_buffer(x = temp_occs, 
                            dist = 10000) %>%  # choose a buffer distance. I chose a 10km radius.
    st_union() %>% 
    st_make_valid() %>% 
    st_intersection(st_make_valid(wm)) %>% 
    st_union()

temp_range <- st_sf(geometry = temp_range) %>% 
  mutate(scientificName_harmonise = i)


for (i in temp_spp[2:length(temp_spp)]) {
message(paste0("running species ", i))
  tryCatch({
temp_occs <- temp_occs_data %>% 
      filter(scientificName_harmonise == i) %>% 
      st_as_sf(coords = c("decimallongitude", "decimallatitude"),
                          crs = "+proj=longlat +datum=WGS84")
temp_range_l <- st_buffer(x = temp_occs, 
                            dist = 10000) %>% 
    st_union() %>% 
    st_make_valid() %>% 
    st_intersection(st_make_valid(wm))

temp_range_l <- st_sf(geometry = temp_range_l) %>% 
  mutate(scientificName_harmonise = i)

temp_range <- rbind(temp_range, temp_range_l)
},
 error = function(e){
   message(paste0("error occurred for species ", i, ":\n"), e)
 })
}

temp_range <- temp_range %>% 
  rename("scientificName_harmonise")

# update the rangemaps

rangeMaps_more <- temp_range %>% 
  left_join(
    temp_occs_data %>% 
    dplyr::select(scientificName_harmonise,
                  family) %>% 
    distinct()
    )

temp  <- st_area(rangeMaps_more)

rangeMaps_more <- rangeMaps_more %>% 
  bind_cols(data.frame(area = as.double(temp)))


rangeMaps <- bind_rows(rangeMaps %>%  rename("area" = "SHAPE_Area"),
                             rangeMaps_more) %>% 
  st_make_valid()

# write data with only geographic range in "./input/" folder

write_csv(st_drop_geometry(rangeMaps) %>% 
            select(scientificName_harmonise, area),
          "./input/geographicRange.csv")

# write the new range map in the input folder
write_sf(rangeMaps,
         "./input/rangeMap.shp")

```

## Human Footprint Maps

Download human footprint raster from: https://sedac.ciesin.columbia.edu/data/set/wildareas-v3-2009-human-footprint
Keep this in "./data/humanFootprint" folder
```{r}
# Read the raster file containing human footprint data for 2009
HF_2009 <- rast(file.choose())

# Set values greater than 50 to NA (assuming these are units indicating human impact)
HF_2009[HF_2009 > 50] <- NA

# Define a new raster with the same extent, resolution of 5000 units, and CRS as the original HF_2009 raster
temp_r <- rast(extent = ext(HF_2009), resolution = 5000, crs = crs(HF_2009))

# Resample the HF_2009 raster to match the resolution and extent of the new raster 'temp_r'
HF_2009 <- resample(HF_2009, temp_r)

# Project the HF_2009 raster to WGS84 coordinate reference system (CRS)
HF_2009 <- project(HF_2009, "+proj=longlat +datum=WGS84")

```



## ecoregions

Download shapefiles of ecoregions from Dinerstein et al. 2017: https://academic.oup.com/bioscience/article/67/6/534/3102935#supplementary-data
Keep this file in "./data/ecoregions" folder

```{r}
ecoregions <- st_read(file.choose()) %>% 
  st_make_valid()
```



### habitat shapefiles

Download raster of ecoregions from Jung et al. 2020 : https://www.nature.com/articles/s41597-020-00599-8 
Keep this file in the "./data/habtiats" folder

```{r}
habitat_lvl1 <- rast(file.choose())
```


# Commonness

## Commonness raw


```{r}
# Download the occurrence data from GBIF. Upload the occurrence data. Put the occurrence data in the "./data/occurrenceData" folder.

occs <- read.delim(file = file.choose(),
                            header = TRUE,
                            sep = "\t") %>% 
  rename("scientificName" = "species")
  left_join(occs_spp_resolve,
            by = c("scientificName" = "binomial"))%>%    
  relocate(scientificName_harmonise)

# Compute the number of records per species. This only include species with geographic coordinates.

commonness_raw <- occs %>% 
 dplyr::select(scientificName_harmonise,
                decimalLongitude,
                decimalLatitude,
                countryCode,
                individualCount,
                gbifID,
                family, 
                taxonRank, 
                coordinateUncertaintyInMeters, 
                year,
                basisOfRecord, 
                institutionCode) %>% 
  rename("decimallongitude" = "decimalLongitude",
         "decimallatitude" = "decimalLatitude") %>%
  mutate_all(na_if, "") %>% 
  mutate(decimallongitude = as.numeric(decimallongitude),
         decimallatitude = as.numeric(decimallatitude)) %>% 
  filter(!is.na(scientificName_harmonise),
         !is.na(decimallongitude),
         !is.na(decimallatitude)) %>% 
  group_by(scientificName_harmonise) %>% 
  summarise(commonness_rawGBIF = n())

```

## Commonness cleaned

```{r}
# Compute the number of high-quality records.

commonness_cleaned <- occs %>%
  dplyr::select(scientificName_harmonise,
                decimalLongitude,
                decimalLatitude,
                countryCode,
                individualCount,
                gbifID,
                family, 
                taxonRank, 
                coordinateUncertaintyInMeters, 
                year,
                basisOfRecord, 
                institutionCode) %>% 
  rename("decimallongitude" = "decimalLongitude",
         "decimallatitude" = "decimalLatitude") %>%
  mutate_all(na_if, "") %>% 
  mutate(decimallongitude = as.numeric(decimallongitude),
         decimallatitude = as.numeric(decimallatitude)) %>% 
  filter(!is.na(scientificName_harmonise),
         !is.na(decimallongitude),
         !is.na(decimallatitude)) %>% 
  cc_dupl()  %>% 
  cc_zero() %>% 
  cc_equ() %>% 
  cc_val() %>% 
  cc_sea() %>% 
  cc_cap(geod = FALSE, buffer = 0.0416) %>% 
  cc_cen(geod = FALSE, buffer = 0.0416) %>% 
  cc_gbif(geod = FALSE, buffer = 0.0416) %>% 
  cc_inst(geod = FALSE, buffer = 0.0416) %>% 
  cc_iucn(range = rangeMaps) %>%
  filter(coordinateUncertaintyInMeters  < 5000  | is.na(coordinateUncertaintyInMeters)) %>% 
  group_by(scientificName_harmonise) %>% 
  summarise(commonness_cleanedGBIF = n())

```

## Commonness thinned 


```{r}
# Compute the number of high-quality, geographically thinned records.
# Geographic thinning reduces sampling bias/sampling pseudo-replicates.

# Extract unique species names from the 'commonness_cleaned' dataframe
temp <- commonness_cleaned %>% 
  distinct(scientificName_harmonise) %>%          # Extract unique values of 'scientificName_harmonise'
  arrange(scientificName_harmonise) %>%           # Arrange them in alphabetical order
  pull(scientificName_harmonise)                  # Pull the 'species' column as a vector

# Initialize an empty dataframe to store the results
commonness_thinned <- data.frame()

# Loop through each species
for (i in temp) {
  tryCatch(
    {
      # Filter data for the current species
      i_sf <- commonness_cleaned %>% 
        filter(scientificName_harmonise == i)
      
      # Convert spatial data to Simple Features
      i_vect <- vect(i_sf,
                     geom = c("decimallongitude", "decimallatitude"),
                     crs = "+proj=longlat +datum=WGS84")
      
      # Convert Simple Features to raster
      i_rast <- rast(i_vect)
      
      # Set resolution of the raster
      res(i_rast) <- 0.0416
      
      # Extend the raster by 1 cell
      i_rast  <- extend(i_rast, ext(i_rast) + 1)
      
      # Perform spatial thinning
      i_thinned <- spatSample(i_vect,
                              size = 1, 
                              "random", 
                              strata = i_rast) 
      
      # Summarize the thinned data and store it
      commonness_thinned <- as.data.frame(i_thinned) %>%
        group_by(scientificName_harmonise) %>% 
        summarise(commonness_thinnedGBIF = n()) %>% 
        rbind(commonness_thinned)
    },
    error = function(e){
      message(paste0("Error occurred for species ", i, ":\n"), e)
    }
  )
}


```

## the compiled commonness data

```{r}
commonness_df <- full_join(commonness_raw,
                           commonness_cleaned) %>% 
  full_join(commonness_thinned)

# write this in input folder
commonness_df %>% 
  write_csv("./input/commonness.csv")
```




# Tolerance to human disturbance



## humanTolerance_range


```{r}
## Compute human tolerance based on entire geographic range

# Extract unique harmonized scientific names from the 'rangeMaps' dataframe
temp_spp <- rangeMaps %>%
  st_drop_geometry() %>%           # Drop geometry (spatial) information
  pull(scientificName_harmonise) %>%  # Extract 'scientificName_harmonise' column
  unique()                            # Keep only unique values

# Initialize an empty dataframe to store the results
humanTolerance_rast <- data.frame()

# Loop through each species
for (i in temp_spp) {
  message(paste0("Running species ", i))
  tryCatch(
    {
      # Filter rangeMaps for the current species
      temp_range <- rangeMaps %>% 
                    filter(scientificName_harmonise == i)
      
      # Mask HF_2009 raster with the range of the current species
      temp_rast <- mask(HF_2009, temp_range)
      
      # Compute statistics (mean, median, range) for the masked raster
      humanTolerance_rast <- cbind(global(temp_rast, fun = "mean", na.rm = TRUE),
                                    global(temp_rast, fun = median, na.rm = TRUE),
                                    global(temp_rast, fun = "range", na.rm = TRUE)) %>% 
        rowid_to_column() %>%   # Add a column for row index
        mutate(scientificName_harmonise = i) %>%   # Add a column for species name
        dplyr::select(scientificName_harmonise,
                      humanTolerance_rast_mean = mean, 
                      humanTolerance_rast_median = global, 
                      humanTolerance_rast_min = range, 
                      humanTolerance_rast_max = max) %>%   # Select specific columns
        # Round numeric columns to 2 decimal places
        mutate(humanTolerance_rast_mean = round(humanTolerance_rast_mean, 2),
               humanTolerance_rast_median = round(humanTolerance_rast_median, 2),
               humanTolerance_rast_min = round(humanTolerance_rast_min, 2),
               humanTolerance_rast_max = round(humanTolerance_rast_max, 2)) %>% 
        rbind(humanTolerance_rast)   # Append results to the main dataframe
    },
    error = function(e){
      message(paste0("Error occurred for species ", i, ":\n"), e)
    }
  )
}


```


## Computing human tolerance on occurrence locations recorded from 2009 to present


# humanTolerance_occs

```{r}
# Download the occurrence data from GBIF. Upload the occurrence data. Put the occurrence data in the "./data/occurrenceData" folder.

occs <- read.delim(file = file.choose(),
                            header = TRUE,
                            sep = "\t") %>% 
  rename("scientificName" = "species")

# Clean the occurrence data.

occs <- occs %>%
  dplyr::select(scientificName,
                decimalLongitude,
                decimalLatitude,
                countryCode,
                individualCount,
                gbifID,
                family, 
                taxonRank, 
                coordinateUncertaintyInMeters, 
                year,
                basisOfRecord, 
                institutionCode) %>% 
  rename("decimallongitude" = "decimalLongitude",
         "decimallatitude" = "decimalLatitude") %>%
  mutate_all(na_if, "") %>% 
  mutate(decimallongitude = as.numeric(decimallongitude),
         decimallatitude = as.numeric(decimallatitude)) %>% 
  filter(!is.na(species),
         !is.na(decimallongitude),
         !is.na(decimallatitude)) %>% 
  cc_dupl()  %>% 
  cc_zero() %>% 
  cc_equ() %>% 
  cc_val() %>% 
  cc_sea() %>% 
  cc_cap(geod = FALSE, buffer = 0.0416) %>% 
  cc_cen(geod = FALSE, buffer = 0.0416) %>% 
  cc_gbif(geod = FALSE, buffer = 0.0416) %>% 
  cc_inst(geod = FALSE, buffer = 0.0416) %>% 
  cc_iucn(range = rangeMaps) %>%
  filter(coordinateUncertaintyInMeters  < 5000  | is.na(coordinateUncertaintyInMeters)) %>% 
  left_join(occs_spp_resolve,
            by = c("scientificName" = "binomial"))%>%    
  relocate(scientificName_harmonise)



temp_spp <- occs %>% 
  distinct(scientificName_harmonise) %>%
  arrange(scientificName_harmonise) %>% 
  pull(scientificName_harmonise)

# Computing human tolerance from occurrence locations

# Initialize an empty dataframe to store the results
humanTolerance_occs <- data.frame()

# Loop through each species
for (i in temp_spp) {
  tryCatch(
    {
      message(paste0("Processing ", i))
      
      # Filter occurrence data for the current species since 2009
      temp_sf <- occs %>% 
        filter(scientificName == i,
               year >= 2009)
      
      # Convert occurrence data to Simple Features
      temp_vect <- vect(temp_sf,
                        geom = c("decimallongitude", "decimallatitude"),
                        crs = "+proj=longlat +datum=WGS84")
      
      # Convert Simple Features to raster
      temp_rast <- rast(temp_vect)
      
      # Set resolution of the raster
      res(temp_rast) <- 0.0416
      
      # Extend the raster by 1 cell
      temp_rast  <- extend(temp_rast, ext(temp_rast) + 1)
      
      # Thinning occurrence data to a 5km radius distance
      temp_thinned <- spatSample(temp_vect,
                                  size = 1, 
                                  "random", 
                                  strata = temp_rast)
      
      # Extract human footprint data based on thinned occurrence locations
      humanTolerance_occs <- terra::extract(HF_2009, temp_thinned) %>% 
        as.data.frame() %>%
        summarise(humanTolerance_occs_mean = round(mean(HFP2009_int, na.rm = TRUE), 2),
                  humanTolerance_occs_median = round(median(HFP2009_int, na.rm = TRUE), 2),
                  humanTolerance_occs_min = round(min(HFP2009_int, na.rm = TRUE), 2),
                  humanTolerance_occs_max = round(max(HFP2009_int, na.rm = TRUE), 2)) %>% 
        mutate(scientificName_harmonise = i) %>% 
        relocate(scientificName_harmonise) %>% 
        rbind(humanTolerance_occs)
    },
    error = function(e){
      message(paste0("Error occurred for species ", i, ":\n"), e)
    }
  )
}

# Replace certain values with NA
humanTolerance_occs <- humanTolerance_occs %>% 
  replace_with_na_all(condition = ~.x %in% c("N/A", "-Inf", "Inf")) %>% 
  mutate_all(function(x) ifelse(is.nan(x), NA, x))

```


## write the csv in the "./input/" folder

```{r}
# combine the human tolerance data from rasters and occurences

humanTolerance <- full_join(humanTolerance_rast, humanTolerance_occs, by = c("species")) %>% 
  replace_with_na_all(condition = ~.x %in% c("N/A", "-Inf", "Inf")) %>% 
  mutate_all(function(x) ifelse(is.nan(x), NA, x))


write_csv(humanTolerance, "./input/humanTolerance.csv")
```



# Access to ports

## loading data

### all transport ports

You can download the data from The World Banks : https://datacatalog.worldbank.org/search/dataset/0038118/Global---International-Ports
Keep file in "./data/ports" folder

```{r}
# Read the CSV file containing port data and convert it to a spatial dataframe
allPorts <- st_as_sf(read_csv(file.choose()),
                      coords = c("X", "Y"),  # Specify the column names for coordinates
                      crs = 4326) %>%       # Specify the coordinate reference system (CRS)
            # Add columns to identify different types of ports based on Function column values
            mutate(
              is_Port = if_else(str_sub(Function, 1, 1) == 1, TRUE, FALSE),
              is_Rail = if_else(str_sub(Function, 2, 2) == 2, TRUE, FALSE),
              is_Road = if_else(str_sub(Function, 3, 3) == 3, TRUE, FALSE),
              is_Airport = if_else(str_sub(Function, 4, 4) == 4, TRUE, FALSE),
              is_Postal = if_else(str_sub(Function, 5, 5) == 5, TRUE, FALSE),
              is_MultimodalFunc = if_else(str_sub(Function, 6, 6) == 6, TRUE, FALSE),
              is_FixedTransportFunc = if_else(str_sub(Function, 7, 7) == 7, TRUE, FALSE),
              is_BorderCrossing = if_else(str_sub(Function, 8, 8) == "B", TRUE, FALSE)
            ) %>%
            st_make_valid()  # Make sure geometries are valid


# Filter allPorts to create separate spatial dataframes for railPorts, roadPorts, portPorts, postalPorts, and borderCrossingPorts
railPorts <- allPorts %>% 
  filter(is_Rail == TRUE) %>%   # Filter for rows where is_Rail is TRUE
  st_make_valid()               # Make sure geometries are valid

roadPorts <- allPorts %>% 
  filter(is_Road == TRUE) %>%   # Filter for rows where is_Road is TRUE
  st_make_valid()               # Make sure geometries are valid

portPorts <- allPorts %>% 
  filter(is_Port == TRUE) %>%   # Filter for rows where is_Port is TRUE
  st_make_valid()               # Make sure geometries are valid

postalPorts <- allPorts %>% 
  filter(is_Postal == TRUE) %>% # Filter for rows where is_Postal is TRUE
  st_make_valid()               # Make sure geometries are valid

borderCrossingPorts <- allPorts %>% 
  filter(is_BorderCrossing == TRUE) %>%  # Filter for rows where is_BorderCrossing is TRUE
  st_make_valid()                        # Make sure geometries are valid

  
```


#### airports only

You can download the data from The World Banks : https://datacatalog.worldbank.org/search/dataset/0038117/Global-Airports
Keep filw in "./data/ports" folder

```{r}
# Read the CSV file containing airport data and convert it to a sf object
airPorts <- st_as_sf(read_csv(file.choose()),   # Read the CSV file
                     coords = c("Airport1Longitude", "Airport1Latitude"),   # Specify the columns containing longitude and latitude
                     crs = 4326) %>%   # Specify the coordinate reference system (CRS) as WGS84 (EPSG code 4326)
  st_make_valid()   # Make the spatial object valid (fix any invalid geometries)

# The airport data is now stored in 'airPorts' as an sf object

```



## Computing ports within range


```{r}
## Computing ports within range

# Extract unique harmonized scientific names from the 'rangeMaps' dataframe
temp_spp <- rangeMaps %>%
  pull(scientificName_harmonise) 

# Define a function to count ports within range for a given species
countPorts <- function(x){
  tryCatch({
    # Filter rangeMaps for the current species and make the geometry valid
    temp_range <- rangeMaps %>% filter(scientificName_harmonise == x) %>% st_make_valid()
    
    # Initialize a vector to store port counts
    temp_res <- rep(NA, 6)
    
    # Calculate the number of intersecting ports for different types of ports
    temp_res[1] <- sum(unlist(st_intersects(airPorts, temp_range)))
    temp_res[2] <- sum(unlist(st_intersects(allPorts, temp_range)))
    temp_res[3] <- sum(unlist(st_intersects(railPorts, temp_range)))
    temp_res[4] <- sum(unlist(st_intersects(roadPorts, temp_range)))
    temp_res[5] <- sum(unlist(st_intersects(postalPorts, temp_range)))
    temp_res[6] <- sum(unlist(st_intersects(borderCrossingPorts, temp_range)))
    
    # Return the result
    return(temp_res)
  },
  error = function(e){
    return(temp_res) # Return NA in case of error
  },
  warning = function(w){
    return(temp_res) # Return NA in case of warning
  }
 )
}

# Initialize a cluster with 6 workers for parallel processing
cl <- makeCluster(6)

# Load required libraries into the cluster workers
clusterEvalQ(cl, 
             {library(dplyr)
              library(sf)})
              
# Export necessary objects to the cluster workers
clusterExport(cl, c("countPorts", 
                    "rangeMaps",
                    "airPorts",
                    "allPorts",
                    "railPorts",
                    "roadPorts",
                    "postalPorts",
                    "borderCrossingPorts"))

# Apply the countPorts function in parallel to each species
temp_res <- clusterApplyLB(cl, temp_spp, countPorts)

# Stop the cluster to release system resources
stopCluster(cl)

# Create a dataframe from the results and write it to a CSV file
data.frame(scientificName_harmonise = temp_spp,
           do.call(rbind, temp_res)) %>% 
  rename("airport_range" = "X1",
         "allPorts_range" = "X2",
         "railPorts_range" = "X3",
         "roadPorts_range" = "X4",
         "postalPorts_range" = "X5",
         "borderCrossingPorts_range" = "X6") %>% 
  distinct(scientificName_harmonise, .keep_all = TRUE) %>% 
  write_csv("./input/portsIntercepted.csv")


```


# Habitat generalism

the process first involves generating a grid of 5km x 5km cells based on the geographic ranges of the species

```{r}
spp_list <- rangeMaps %>%  
  st_drop_geometry() %>% 
  arrange(scientificName_harmonise) %>% 
  distinct(scientificName_harmonise) %>% 
  pull(scientificName_harmonise) 


habitatGeneralism_range <- function(x) {
  # Extract the species name for the current iteration
  temp_spp <- spp_list[x]
  
  # Initialize a vector to store the results
  temp_res <- rep(NA, 5)
  
  tryCatch({
    # Attempt to execute the following code block
    
    # Filter rangeMaps for the current species
    temp_range <- rangeMaps %>% 
      filter(scientificName_harmonise == temp_spp)
    
    # Extract available habitats represented by ecoregions
    temp_ecoregions <- ecoregions[temp_range,] %>% 
      st_make_valid()
    
    # Extract occupied habitats (among available) by cropping habitat_lvl1 with the species range
    temp_habitat <- terra::crop(habitat_lvl1, vect(temp_range), mask = TRUE) %>%
      terra::mask(temp_range)
    
    # Extract available habitats by cropping habitat_lvl1 with the available ecoregions
    temp_available <- terra::crop(habitat_lvl1, vect(temp_ecoregions), mask = TRUE) %>%
      terra::mask(temp_ecoregions)
    
    # Create a frequency table for occupied and available habitats
    temp_table <- full_join(
      freq(temp_habitat) %>% 
        mutate(resource = value) %>% 
        dplyr::select(-c(value)) %>% 
        rename("occupied" = "count"),
      freq(temp_available) %>% 
        mutate(resource = value) %>% 
        dplyr::select(-c(value)) %>% 
        rename("available" = "count"),
      by = c("layer", "resource")
    ) %>% 
      mutate(
        occupied = replace_na(occupied, 0),
        prop.Occupied = round(occupied / sum(occupied), 4),
        prop.Available = round(available / sum(available), 4),
        propDiff = abs(prop.Occupied - prop.Available)
      )
    
    # Calculate the most used habitat and round it to 4 decimal places
    temp_res[1] <- temp_table %>% 
      arrange(desc(occupied)) %>%
      head(1) %>% 
      pull(resource) %>% 
      round(4)
    
    # Calculate habitat breadth (number of habitats occupied) and round it to 4 decimal places
    temp_res[2] <- temp_table %>% 
      summarise(sum(occupied != 0)) %>%
      pull() %>% 
      round(4)
    
    # Calculate Levins' measure of niche breadth and round it to 4 decimal places
    temp_res[3] <- temp_table %>% 
      summarise(1 / sum(prop.Occupied^2)) %>%
      pull() %>% 
      round(4)
    
    # Calculate normalized Levins' measure of niche breadth and round it to 4 decimal places
    temp_res[4] <- temp_table %>% 
      summarise(1 / (n() * sum(prop.Occupied^2))) %>%
      pull() %>% 
      round(4)
    
    # Calculate proportional similarity index and round it to 4 decimal places
    temp_res[5] <- temp_table %>% 
      summarise(1 - (0.5 * sum(propDiff))) %>%
      pull() %>% 
      round(4)
    
    # Create a list containing the species name and the calculated results
    temp_res_list <- list(temp_spp, temp_res)
    
    # Return the list
    return(temp_res_list)
  },
  error = function(e) {
    # Handle errors by returning NA values
    temp_res_list <- list(temp_spp, temp_res)
    return(temp_res_list)
  },
  warning = function(w) {
    # Handle warnings by returning NA values
    temp_res_list <- list(temp_spp, temp_res)
    return(temp_res_list)
  })
}


# Setting up a cluster with 6 workers for parallel processing
cl <- makeCluster(6)  # You may adjust the number of workers based on the available CPUs of your computer

# Loading required libraries into the cluster workers
clusterEvalQ(cl, {
  library(dplyr)   # For data manipulation
  library(tidyr)   # For data tidying
  library(sf)      # For spatial data handling
  library(terra)   # For raster data manipulation
})

# Exporting necessary objects to the cluster workers
clusterExport(cl, c("rangeMaps",   # Assuming these are objects or functions needed for the computation
                    "ecoregions",
                    "spp_list",
                    "habitatGeneralism_range"))

# Applying a function called habitatGeneralism_range in parallel to each element of the spp_list
# The function habitatGeneralism_range is applied across the workers using load balancing
temp_res <- clusterApplyLB(cl, seq(1:length(spp_list)), habitatGeneralism_range)

# Stopping the cluster to release system resources
stopCluster(cl)



### save data in  folder

data.frame(scientificName_harmonise = unlist(purrr::map(temp_res, 1)),
           do.call(rbind, purrr::map(temp_res, 2))) %>% 
  rename("habitat_mostFreq" = "X1",
         "habitatBreadth" = "X2",
         "Levins" = "X3",
         "Levins_norm" = "X4",
         "PropSim" = "X5") %>% 
  write_csv(".//habitatGeneralism.csv")
```
